---
title: "Prediction of the manner in which the exercise was done"
author: "Prabhu Manickavelu"
date: "2 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Question and about the data
We are going to try predict how a person did the exercise.

There the data consistes of 5 type of doing the same exercise. which is classified is **classe** variable. The following are the classified

* A:- Exactly according to the specification (The correct fashion).
* B:- Throwing the elbows to the front (The wrong fashion).
* C:- Lifting the dumbbell only halfway (The wrong fashion).
* D:- Lowering the dumbbell only halfway (The wrong fashion).
* E:- Throwing the hips to the front (The wrong fashion).

### Note
We are not trying to identfy which is the correct fashion. but the fashion used by the person. 

```{r warning=FALSE, message=FALSE}
library(caret)
library(rpart)
library(randomForest)
```

## Getting and reading the training and test data.
```{r cache=TRUE}
training <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"), na.strings=c("NA","#DIV/0!",""))

quize <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"), na.strings=c("NA","#DIV/0!",""))
num_rows <- nrow(training)
```
Number of rows of data in training set: **`r num_rows`**
```{r eval=FALSE}
summary(training) # The summary is shown in the Annex 
```
By Checking the summary there are may columns with lots of *N/A* values

## Cleaning the data by removeing columns with over *30% of N/A values* and unnecessary columns.
```{r cache=TRUE}
# Getting all column names
col_names <- colnames(training) 
# Removing the known unwanted columns.
col_names <- col_names[!col_names %in% c("X", "user_name", "new_window")]
clean_training <- data.frame(training[,col_names])
# Looping through the columns to find columns with over 30% of N/A values and removing them
for (col_name in length(col_names):1) {
    if(sum(is.na(clean_training[,col_name]))/num_rows >= .7){
        clean_training <- clean_training[,-col_name]
    }
}
```

## Spliting the data for training and test data 70% training and 30% testing
```{r cache=TRUE}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
clean_training_ds <- clean_training[inTrain,]
clean_testing_ds <- clean_training[-inTrain,]
```

## Trying out Diffrent algorithms.
### RPart
```{r}
# Training the data
rpart_fit <- train(classe ~ ., data = clean_training_ds, method="rpart")
# predecting the test data
rpart_predict <- predict(rpart_fit, clean_testing_ds[,-58])
confusionMatrix(rpart_predict, clean_testing_ds$classe)
```

### RandomForest
```{r cache=TRUE}
# Training the data
rf_fit <- train(classe ~ ., data = clean_training_ds, method="rf")
# predecting the test data
rf_predict <- predict(rf_fit, clean_testing_ds[,-58])
confusionMatrix(rf_predict, clean_testing_ds$classe)
```

## Annex
### Summary of the raw training data.
```{r}
summary(training) # The summary is shown in the Annex 
```
